Efficiently Handling Guards in the Low Level Design of RPython's Tracing JIT

### 2012-08-07 talk/vmil2012/paper.tex
keywords and another todo
(cfbolz, bivab) write about bridges and trace stitching

\keywords{XXX}

\begin{abstract}
In pellentesque faucibus vestibulum. Nulla at nulla justo, eget luctus tortor.
Nulla facilisi. Duis aliquet egestas purus in blandit. Curabitur vulputate,
ligula lacinia scelerisque tempor, lacus lacus ornare ante, ac egestas est urna
sit amet arcu. Class aptent taciti sociosqu ad litora torquent per conubia
nostra, per inceptos himenaeos. Sed molestie augue sit amet leo consequat
posuere. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices
posuere cubilia Curae; Proin vel ante a orci tempus eleifend ut et magna. Lorem
ipsum dolor sit amet, consectetur adipiscing elit. Vivamus luctus urna sed urna
ultricies ac tempor dui sagittis. In.
\end{abstract}

@@

If a guard fails often enough, a trace is started from it
to create a trace tree.
When that happens another use case of resume data
is to construct the tracer state.

@@

% subsection Interaction With Optimization (end)
\subsection{Compiling Side-Exits and Trace Stitching} % (fold)
\label{sub:Compiling side-exits and trace stitching}
   * tracing and attaching bridges and throwing away resume data
   * restoring the state of the tracer
     * keeping virtuals
   * compiling bridges
\todo{maybe mention that the failargs also go into the bridge}

% subsection Compiling side-exits and trace stitching (end)
% section Resume Data (end)

@@

Once the bridge has been compiled the guard that led to compiling the bridge is
patched to redirect control flow to the bridge in case the check fails. In
future if the guard fails again it jumps to the code compiled for the bridge
instead of bailing out. Once the guard has been compiled and attached to the
loop the guard becomes just a point where control-flow can split. The loop
after the guard and the bridge are just conditional paths.
Figure~\ref{fig:trampoline} shows a digram of a compiled loop with two guards,
Guard \#1 jumps to the trampoline, loads the \texttt{low level resume data} and
then calls the compensation code, whereas Guard \#2 has already been patched
and directly jumps to the corresponding bridge. The bridge also contains two
guards that work based on the same principles.
\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/loop_bridge.pdf}
\caption{Trace control flow in case of guard failures with and without bridges}
\label{fig:trampoline}
\end{figure}
%* Low level handling of guards
%   * Fast guard checks v/s memory usage
%   * memory efficient encoding of low level resume data
%   * fast checks for guard conditions
%   * slow bail out
%
% section Guards in the Backend (end)

>>>>

\keywords{tracing JIT, guards, deoptimization}

\begin{abstract}
\todo{write}
In pellentesque faucibus vestibulum. Nulla at nulla justo, eget luctus tortor.
Nulla facilisi. Duis aliquet egestas purus in blandit. Curabitur vulputate,
ligula lacinia scelerisque tempor, lacus lacus ornare ante, ac egestas est urna
sit amet arcu. Class aptent taciti sociosqu ad litora torquent per conubia
nostra, per inceptos himenaeos. Sed molestie augue sit amet leo consequat
posuere. Vestibulum ante ipsum primis in faucibus orci luctus et ultrices
posuere cubilia Curae; Proin vel ante a orci tempus eleifend ut et magna. Lorem
ipsum dolor sit amet, consectetur adipiscing elit. Vivamus luctus urna sed urna
ultricies ac tempor dui sagittis. In.
\end{abstract}

@@

If a guard fails often enough, a trace is started from it
to create a trace tree.
When that happens another use case of resume data
is to construct the tracer state.
After the bridge has been recorded and compiled it is attached to the guard.
If the guard fails later, the bridge is executed. Therefore the resume data of
that guard is no longer needed.

@@

% section Resume Data (end)

@@

Once the bridge has been compiled the guard that led to compiling the bridge is
patched to redirect control flow to the bridge in case the check fails. In
future if the guard fails again it jumps to the code compiled for the bridge
instead of bailing out. Once the guard has been compiled and attached to the
loop the guard becomes just a point where control-flow can split. The loop
after the guard and the bridge are just conditional paths.
Figure~\ref{fig:trampoline} shows a digram of a compiled loop with two guards,
Guard \#1 jumps to the trampoline, loads the \texttt{low level resume data} and
then calls the compensation code, whereas Guard \#2 has already been patched
and directly jumps to the corresponding bridge. The bridge also contains two
guards that work based on the same principles.
\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/loop_bridge.pdf}
\caption{Trace control flow in case of guard failures with and without bridges}
\label{fig:trampoline}
\end{figure}
% section Guards in the Backend (end)
